---
title: "Average Real Hourly Earnings for the Education and Training Sector of New Zealand \nQuarterly GDP for Retail Trade in New Zealand"
fontsize: 11pt
output:
  bookdown::html_document2:
    fig_height: 5
    toc: yes
    toc_depth: 1
    toc_float:
      collapsed: false
    number_sections: false
    code_folding: show
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fpp3)
library(kableExtra)
```




# Average Real Hourly Earnings for the Education and Training Sector of New Zealand

We have quarterly time series data on average hourly earnings in the education and training sector of New Zealand, from `2012Q1` until `2022Q4`.  Nominal earnings are measured in the prices of the day and real earnings are measured in the prices of a fixed base date.  The file `earnings.csv` contains average nominal hourly earnings for the education and training sector and the file `cpi.csv` contains the Consumer Price Index (CPI) with base-quarter 2022 Q4.  The following code reads these data sets into `R`. 

```{r, message = FALSE}
earnings <- read_csv("earnings.csv") %>% 
  mutate(Quarter = yearquarter(Quarter)) %>%
  as_tsibble(index = Quarter)
earnings

cpi <- read_csv("cpi.csv") %>% 
  mutate(Quarter = yearquarter(Quarter)) %>%
  as_tsibble(index = Quarter)
cpi
```


```{r}
earnings %>% 
  slice (1 : 10)%>%
  knitr::kable()

cpi %>% 
  slice (1 : 10)%>%
  knitr::kable()
```



```{r}

joined_data <- earnings %>%
  left_join(cpi, by = "Quarter") 

joined_data

```



```{r}

base <- cpi %>% 
  pull()  %>% 
  last


adjusted_data <- joined_data %>% 
  mutate(Real =  Nominal / CPI * base) %>%
  select(-Nominal, -CPI) #I had to remove these 2 because it wouldn't plot `real`
adjusted_data


```


```{r}

fit <- adjusted_data  %>%
  model(SES = ETS(Real ~ error("A") + trend("N") + season("N")),
        HL = ETS(Real ~ error("A") + trend("A") + season("N")),
        HDL = ETS(Real ~ error("A") + trend("Ad") + season("N")))
fit
```



```{r}
adjusted_data %>%
  autoplot(mapping = aes(x = Quarter,
                         y = Real)) + 
  geom_line(data = augment(fit), 
            aes(y = .fitted, 
                colour = .model)) + 
  theme_minimal()
```



```{r}
report(fit) %>%
  select(.model,AICc)

```

The Akaike Information Criterion with Correction (AICc) asseess the goodness of fit of statistical models. Models with lower AICc values are considered better fits to the data. In this example, the Simple Exponential Smoothing (SES) method was found to have the lowest AICc value, indicating that it provides the best fit for the data. The Holts Damped Linear and Holts Linear models were found to have larger AIC values, suggesting that these models are overfitting the data. This could be linked to their complexity, as they may be capturing statistical noise instead of true patterns in the data. 





```{r}

fit %>%
  select(SES) %>%
  report()
```
`alpha`{.bg-info}
- The output shows that Alpha is 0.95, which indicates that the Simple Exponential Smoothing model (SES) places more weight on recent observations and less weight on past observations when forecasting the Real variable.


`Extra: sigma^2`{.bg-info}
The formula for AIC is AIC = -2ln(L) + 2k, where L is the likelihood of the model and k is the no. of parameters and initial states estimated (including the residual variance. Therefore, a larger value of sigma^2 will increase the penalty in the AIC calculation, and favor simpler models with lower residual variance.




# Quarterly GDP for Retail Trade in New Zealand

We have quarterly (chain volume) gross domestic product (QGDP) for the retail trade industry in New Zealand, measured in the prices from 2009/10 in NZD Millions. The data, courtesy of Statistics New Zealand, contains the actual time series as well as the seasonally-adjusted time series, measured from `1987Q2` until `2022Q4`. The following code reads this data set into `R`. 

```{r, message = FALSE}
gdp <- read_csv("qgdp_retail.csv") %>%
  mutate(Date = yearquarter(Date)) %>%
  as_tsibble(index = Date)

gdp%>% 
  autoplot()
```


```{r}
dummy1 <- gdp %>%
  mutate(outlier = ifelse(Date %in% yearquarter("2020 Q2"), 1, 0))


gdp_data <- dummy1 %>%
  mutate(unusual = ifelse(Date %in% yearquarter("1989 Q2"), 1, 0))

gdp_data

```



```{r}

fit <- gdp_data  %>%
  model( m1 = TSLM(log(Actual) ~ trend() + season()),
         m2 = TSLM(log(Actual) ~ trend() + season() + outlier + unusual),
         m3 = TSLM(log(Actual) ~ trend(knots = (yearquarter(c("2008 Q1", "2010 Q1")))) + season() + outlier + unusual)) 
  
fit

         
```



```{r}

glance(fit) %>%
  select(.model, AICc)

# Selecting the best model
fit <- fit %>%
  select(m3)

fit %>%
  report()

```


Information criteria, such as AICc, are commonly used for model selection and comparison. Compared to the multiple R^2 for assessing goodness of fit, AICc provides a more robust measure of model performance by penalizing models with many parameters. This is because multiple R^2 can overestimate the goodness of fit for models with too many predictors. Therefore, AICc is considered a better alternative to multiple R^2 for assessing the best fit for the data. 

Therefore, Model 3 is the best model as it incorporates a piecewise linear trend with knots at 2008 Q1 and 2010 Q1 to model the global financial crisis, seasonal dummies, and two outlier dummy variables, has the lowest AICc value among the competing models.



```{r}
# Linearity?
gdp %>%
  autoplot(log(Actual)) + 
  theme_minimal()


# Independence, Normality, Equal Variance?
fit %>%
  gg_tsresiduals()


# Perform Ljung-Box test
fit %>% 
  augment() %>%
  features(.innov, features = ljung_box, lag = 16, dof = 9)
```

`Linearity`{.bg-info}
The plots suggest that there is a linear relationship between the log of actual GDP and time, but some nonlinear behavior is present. Therefore, we cannot confirm the satisfaction of the linearity assumption.

`Independence`{.bg-info}
The autocorrelation function (ACF) plot shows a significant positive autocorrelation, especially lag 1, indicating that there is some dependence between the current observation and past observations. The Ljung-Box test further proves this finding, as it displays a large statistic value of 355.3609 contributing to a very low P-value of 0. This indicates that we can reject the null hypothesis of white noise, and that there's a relationship between the current time and the previous time. 

`Normality`{.bg-info}
The distribution of the innovation residuals is not normally distributed, as evidenced by the asymmetric distribution caused by an observation located further to the right. Therefore, the right skew defies the normality distribution assumption.

`Equal Variance`{.bg-info}
The residuals plot suggests the presence of heteroscedasticity, as peaks and troughs are present over time. This pattern indicates that the variability of the residuals changes over time and violates the assumption of equal variance.

Overall, the analysis suggests that the model does not satisfy the key assumptions, including linearity, independence, normality, and equal variance. As such, the validity and reliability of the model's results are suspected, and alternative models or adjustments may be required to accurately capture the relationships present in the data.



```{r}
future <- new_data(gdp, n = 8)  

future <- future %>%
  mutate(outlier = rep(c(0, 0, 0, 0), 2),
         unusual = rep(c(0, 0, 0, 0), 2))


# Forecast 8 Quarters ahead
fc <- fit %>%
  select(m3) %>%
  forecast(new_data = future)
fc
```

```{r}
# Plot forecast
fc %>%
  autoplot(gdp, level = 90) + 
  theme_minimal()
```

```{r}
# JUST TO SEE WHAT IT LOOKS LIKE 
# COMPARING THE APPLICATION OF THE MODEL TO THE DATA VS RAW DATA. 
#           !!!!!!!ignore!!!!!!!
fc %>%
  autoplot(gdp) +
  geom_line(data = augment(fit),
            aes(y = .fitted),
            colour = "blue")
```




```{r}
# Automatic PI
fc_90.2 <- fc %>%
  hilo(level = 90) %>%
  select(`90%`) 
fc_90.2

fc_90.2 <- fc %>%
  hilo(level = 90) %>%
  select(`90%`) %>%
  filter(Date == yearquarter("2023 Q1"))
fc_90.2
```
With 90% probability that a future observation will be contained within the prediction interval, we predict that the actual GDP from a cross-sectional study for 2023 Quarter 1 will be somewhere between 3355.205 and 3914.585 NZD Millions. 




```{css}

body {
  background-color: #b3e6ff;
}
.columns {
  display: flex;
}

h1, h2, h4 {
  text-align: center;
  font-weight: 700;
}



```


